---
title: "ESM 204 HW 3"
author: "Hannah Brady and Janelle Gaun"
date: "5/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ggpubr)
library(broom)
library(knitr)
library(kableExtra)
options(scipen=999)
```


# Part 1

*One gallon of gasoline emits 19.6 pounds of CO~2~. Assuming that the interim Social COst of Carbon (SCC) correctly reflects the total social cost of one metric ton of CO~2~, what is the marginal externality cost (MEC) per gallon of gas?*

The Biden Administration's "interim" value of the SCC is $51 per metric ton of CO2.

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Calculate and save MEC of one gallon of gas

# MEC = ($/ metric ton) * (metric tons/lb) * (lb/gal)

MEC <- 51*(1/2204.62)*(19.6)

```

The MEC can be calculated as follows:

MEC~gal~ = \$51/metric ton CO~2~ * (1 ton/ 2204.64 lbs) * (19.6 lb CO~2~/gal of gas) = \$`r round(MEC,2)` per gallon of gas

# Part 2

*What is the aggregate daily demand curve for gasoline? What is the supply curve for gasoline? What is the "benefit" to consumers under the status quo? What is the "benefit" to producers under the status quo? What is the environmental cost under the status quo?*

## Aggregate daily demand curve

Data provided includes price (in \$) and quantity (in gallons) estimates of demand per day for "high" and "low" income groups.

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Read in data
data <- read_csv("HW3_data.csv")

# Create 2 linear models
D_low_lm = lm(Price_Dollars ~ Q_Low_Gallons, data = data)
D_high_lm = lm(Price_Dollars ~ Q_High_Gallons, data = data)

# Create plot for reference
ggplot(data = data) +
  #Low Demand
  geom_point(aes(x = Q_Low_Gallons, y = Price_Dollars, color = "Low"), size = 3, alpha = 0.5) +
  geom_smooth(aes(x = Q_Low_Gallons, y = Price_Dollars, color = "Low"), method = "lm", se = F)+
  #High Demand
  geom_point(aes(x = Q_High_Gallons, y = Price_Dollars, color = "High"), size = 3, alpha = 0.5) +
  geom_smooth(aes(x = Q_High_Gallons, y = Price_Dollars, color = "High"), method = "lm", se = F)+
  labs(x = "Quantity gas (Gallons)", y = "Marginal cost ($/gal)")+
  theme_light()

# Save intercepts
int_low <- D_low_lm$coefficients[[1]]
m_low <- D_low_lm$coefficients[[2]]

int_high <- D_high_lm$coefficients[[1]]
m_high <- D_high_lm$coefficients[[2]]

```

Data points are fit with a linear model to create the following estimates of demand curves by income group:


D~low~ = `r round(m_low,6)`Q + `r round(int_low,1)`

D~high~ = `r round(m_high,6)`Q + `r round(int_high,1)`


Aggregate demand curve is created by horizontally summing individual demand curves. This is done by solving each equation for Q, adding both equations together, and re-solving for P. The resulting equation is valid for the range of prices where both high and low consumers have demand. The aggregate demand equals the demand for the high demand consumer group at prices where only the high demand consumer group has demand: 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Solve for intercepts for aggregate demand
m_agg <- 1/(1/(m_low)+1/(m_high))
int_agg <- ((int_low/m_low)+(int_high/m_high))/(1/(m_low)+1/(m_high))

# Create functions for all three demand curves
D_low <- function(Q) {m_low*Q + int_low}
D_high <- function(Q) {m_high*Q + int_high}
D_agg <- function(Q) {m_agg*Q + int_agg}
```


D~agg~ , P < `r round(int_low,1)` = `r round(m_agg,6)`Q + `r round(int_agg,1)`

D~agg~ , P > `r round(int_low,1)` = `r round(m_high,6)`Q + `r round(int_high,1)`

## Supply Curve

We are given that:

+ the current gas price (without any taxes) is \$3.00 per gallon, i.e. P~eq~ = \$3/gal
+ the marginal cost of producing a gallon of gasoline is linear
+ MC has a price-intercept of 0, i.e. at Q = 0 gal, P = \$0/gal

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Solve for Q eq given P = 3:
P_eq <- 3
Q_eq <- (P_eq-int_agg)/m_agg

# Save slope of supply curve:
m_MC <- P_eq/Q_eq

# Create function for supply curve
MC <- function(Q) {m_MC*Q}
```

Solve for Q~eq~ by solving for Q, given D~agg~ = \$3/gal:

Q~eq~ = `r round(Q_eq,0)` gallons of gasoline

Solve for slope: (rise/run) = (P~eq~ - 0)/(Q~eq~ - 0) = `r round(m_MC,6)`

MC = `r round(m_MC,6)`Q


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Graph for reference
ggplot(data.frame(Q = c(0, 800000)), aes(x = Q)) +
  stat_function(fun = D_agg, aes(color = "Aggregate Demand")) +
  stat_function(fun = D_low, aes(color = "Low Demand")) +
  stat_function(fun = D_high, aes(color = "High Demand")) +
  stat_function(fun = MC, aes(color = "Supply")) +
  geom_segment(aes(x = 0, y = P_eq, xend = Q_eq, yend = P_eq, color = "Equilibrium"), linetype = 2)+
  geom_segment(aes(x = Q_eq, y = 0, xend = Q_eq, yend = P_eq, color = "Equilibrium"), linetype = 2)+
  scale_x_continuous(name = "Q (Gallons)") +
  scale_y_continuous(name = "P ($/gal)",limits = c(0, NA)) +
  labs(color = "Curve") +
  theme_minimal()
  
```

## Status quo benefits to producers and consumers

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Calculate consumer surplus

# Solve for quantity at kink
Q_kink <- (int_low-int_agg)/m_agg

# Consumer surplus equals area under high demand curve until kink
SQ_CS_1 <- integrate(D_high, lower = 0, upper = Q_kink)

# plus area under agg demand curve after kink until intercept with supply curve
SQ_CS_2 <- integrate(D_agg, lower = Q_kink, upper = Q_eq)

# minus area from price equals 0 to Peq * Qeq
SQ_CS_total <- SQ_CS_1$value + SQ_CS_2$value - Q_eq*P_eq

# Calculate producer surplus

# Solve for area under supply curve:
SQ_PS_1 <- integrate(MC, lower = 0, upper = Q_eq)

# Subtract from Qeq*Peq:
SQ_PS_total <- Q_eq*P_eq - SQ_PS_1$value
```
Consumer surplus is calculated by integrating under the aggregate demand curve from 0 to Q~eq~ and the subtracting the cost of gasoline (P~eq~*Q~eq~):

Consumer surplus under status quo = \$`r round(SQ_CS_total,0)`.

Producer surplus is calculated by integrating under the supply curve to calculate cost of producing gas and subtracting that from total payment by consumers to producers for gasoline (P~eq~*Q~eq~):

Producer surplus under status quo = \$`r round(SQ_PS_total,0)`.

## Environmental cost under status quo

Under the status quo, the total environmental cost is equal to:

Total Environmental Cost = MEC~gal~ (from Part 1) * Total gallons of gas consumed by society (Q~eq~) = \$`r round(MEC,2)`/gal * `r round(Q_eq,0)` gal of gas = \$`r round(MEC*Q_eq,2)`

# Part 3

*How is the current consumer benefit divided between "high" and "low" income consumers?*

Use the demand curves for the high and low demand to solve for the quantities for each consumer group given the equilibrium price of \$3/gal gas:
```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Low demand:
Q_eq_low <- (P_eq-int_low)/m_low

# High demand:
Q_eq_high <- (P_eq-int_high)/m_high

```

+ Q~[eq low]~ = `r round(Q_eq_low,0)` gallons of gas purchased by low demand consumers

+ Q~[eq high]~ = `r round(Q_eq_high,0)` gallons of gas purchased by high demand consumers

Calculate benefit to each consumer group by integrating under the demand curve from 0 to Q~[eq of group]~ and then subtracting value of P~eq~*Q~[eq of group]~:

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Low demand
CS_low_1 <- integrate(D_low, lower = 0, upper = Q_eq_low)
CS_low_total <- CS_low_1$value - P_eq*Q_eq_low

# High demand
CS_high_1 <- integrate(D_high, lower = 0, upper = Q_eq_high)
CS_high_total <- CS_high_1$value - P_eq*Q_eq_high

```

+ Consumer surplus of low demand group = \$`r round(CS_low_total,0)`

+ Consumer surplus of high demand group = \$`r round(CS_high_total,0)`


# Part 4 

*Derive the optimal gas tax (in \$ per gallon of gasoline) using the interim SCC. Noting that recent research has shown the poor face a disproportionate share of the impacts from climate change, assume that the climate externality is borne entirely by the "low" income group. What would be the effects of this tax on:*


The optimal gas tax would be equal to the marginal external cost (MEC) from Part 1:


MEC = $`r round(MEC,2)` per gallon of gas


#### a. The amount of gasoline produced and consumed

Find the marginal social cost of gasoline by vertically aggregating the marginal supply curve and the marginal external cost curves

    MSC  = `r round(m_MC,6)`Q + `r round(MEC,2)`

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Total Social Cost:
MSC <- function(Q){m_MC*Q + MEC}

#Graph for reference
ggplot(data.frame(Q = c(0, 800000)), aes(x = Q)) +
  stat_function(fun = MC, aes(color = "Marginal Private Cost")) +
  stat_function(fun= MSC, aes(color = "Marginal Social Cost"))+
  stat_function(fun = D_agg, aes(color = "Agg Demand")) +
  scale_x_continuous(name = "Q (Gallons)") +
  scale_y_continuous(name = "P ($/gal)",limits = c(0, NA)) +
  labs(color = "Curve") +
  theme_minimal()

```


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Find new equilibrium quantity
MSC_Q_eq <- (MEC-int_agg)/(m_agg-m_MC)
``` 
    The tax results in a socially optimum equilibrium quantity of `r round(MSC_Q_eq,1)` gallons of gas produced and consumed.

#### b. The price of gasoline

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Use the new equilibrium quantity to find the price
MSC_P_eq <- (m_MC*MSC_Q_eq+ MEC)
```
The new price of gasoline is found by solving the marginal social cost curve for P~[socially optimal]~ given the socially optimal equilibrium quantity calculated in Part 4.a above:

  Price~[socially optimal]~ = `r round(m_MC,6)`*`r round(MSC_Q_eq,1)` + `r round(MEC,2)`= \$ `r round(MSC_P_eq,2)` /gal

#### c. Overall welfare of "high" income consumers

Calculate the welfare of "high" income consumer by determining the consumer surplus at the new equilibrium and subtracting the cost of purchasing gas

+ Solve D~high~ = `r round(m_high,6)`Q + `r round(int_high,2)` for Q, at the socially optimum equilibrium price \$`r round(MSC_P_eq, 2)`/gal determined in Part 4.b above

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Quantity purchased at new socially optimal equilibrium 
MSC_Q_eq_high= (MSC_P_eq-int_high)/(m_high)

# High demand
CS_high_MSC <- integrate(D_high, lower = 0, upper = MSC_Q_eq_high)
CS_high_MSC_total <- CS_high_MSC$value - MSC_P_eq*MSC_Q_eq_high
```

    Q~[SO high]~ = `r round(MSC_Q_eq_high, 2)` gallons

+ Integrate "high" demand curve (`r round(m_high, 6)`Q + `r round(int_high, 2)`) from 0 to `r round(MSC_Q_eq_high, 2)` then subtract the cost of purchasing gas (\$`r round(MSC_P_eq, 2)`/gal*`r round(MSC_Q_eq_high, 2)` gal)

    High demander welfare = \$`r round(CS_high_MSC_total,1)` 

#### d. Overall welfare of "low" income consumers

Calculate the welfare to the low income group by determining the consumer surplus at the socially optimal equilibrium price and quantity. Additionally, because the "low" income group bears the entire cost of the negative externality, the total environmental cost of consuming gas by both income groups must also be subtracted from the low demanders consumer surplus.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#MEC*quantity of gas consumed

TED <- MEC*MSC_Q_eq
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Quantity purchased at new socially optimal equilibrium 
MSC_Q_eq_low= (MSC_P_eq-int_low)/(m_low)

# consumer surplus
CS_low_MSC <- integrate(D_low, lower = 0, upper = MSC_Q_eq_low)
CS_low_MSC_total <- CS_low_MSC$value - MSC_P_eq*MSC_Q_eq_low

#Comment: HRB - maybe we don't need DWL and we just subtract the env damage from the CS_low?

##Find the DWL
#Integrate the MSC from the socially optimal equilibrium to the BAU equilibrium. 
MSC_area <- integrate(MSC, lower = MSC_Q_eq, upper = Q_eq)

#Integrate the aggregate demand curve from the socially optimal equilibrium to the BAU equilibrium 
D_agg_area <- integrate(D_agg, lower = MSC_Q_eq, upper = Q_eq)


## Calculate total welfare 
#Subtract the DWL from the consumer surplus identified above

MSC_welfare_low <- (CS_low_MSC_total-TED)
```

+ Solve D~low~ = `r round(m_low,6)`Q + `r round(int_low,2)` for Q, at the socially optimum equilibrium price \$`r round(MSC_P_eq, 2)`/gal determined in Part 4.b above

    Q~[SO low]~ = `r round(MSC_Q_eq_low, 2)`gallons

+ Integrate "low" demand curve (`r round(m_low, 6)`Q + `r round(int_low, 2)`) from 0 to `r round(MSC_Q_eq_low, 2)` then subtract the cost of purchasing gas (\$`r round(MSC_P_eq,2)`/gal*`r round(MSC_Q_eq_low, 2)` gal) and the total cost of environmental damage from all of society's gas consumption (MECxQ~SO~)

    Low demander welfare = \$`r round(MSC_welfare_low,2)`
 

#### e. Gas producers


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Amount of gas sold 
income <- MSC_P_eq*MSC_Q_eq - TED

#Cost of production
cost <- integrate (MC, lower = 0, upper = MSC_Q_eq)

#Revenue
revenue <- income - cost$value
```

Find the welfare of the producers by finding the total income generated at the socially optimal equilibrium and subtracting the tax revenue and the cost of production. 

+ Income = P~SO~*Q~SO~ - tax revenue
+ Cost: Integrate MPC curve (`r round(m_MC, 6)`Q) from 0 to Q~SO~ (`r round(MSC_Q_eq,0)` gal)
+ Welfare: Income - Cost

Welfare = \$ `r round(revenue, 2)`

#### f. Total environmental damage

Total environmental damage is found by multiplying the MEC (\$`r round(MEC, 2)` per gallon) by the quantity of gas consumed Q~SO~ (`r round(MSC_Q_eq,2)`) gallons



TED = \$`r round(TED,2)`

#### g. Total tax revenue generated


Because the tax equals the MEC, tax revenue equals total environmental damage, which was found by multiplying the MEC/tax (\$`r round(MEC, 2)` per gallon) by the quantity consumed Q~SO~ (`r round(MSC_Q_eq,2)`) gallons

Tax Revenue= \$`r round(TED,2)` 

# Part 5

*Now, assume that all revenue from the gas tax will be redistributed to the consumers in proportion to their pre-tax consumption of gas. For example, if 80% of the gas was consumed by “high” income consumers, then they get 80% of the tax revenue. Additionally, consider the fact that current scientific evidence suggests the true SCC may be much higher than $51. For a range of SCC values ($51, $75, $100, $125, and $150 per metric ton of CO~2~), calculate the effects of an SCC-based gasoline tax on overall welfare of “high” income consumers, overall welfare of “low” income consumers, and gas producers.*

Taxes are redistributed proportional to original consumption of gas between high and low income groups at the free market equilibrium price as determined in Part 3 above.

Proportion= Q~group~/Q~total~

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Proportion of "high"
prop_high <- Q_eq_high/Q_eq
pct_high <- 100*prop_high

#Proportion of "low"
prop_low <- Q_eq_low/Q_eq
pct_low <- 100*prop_low
```
+ Percentage Consumed~low~= `r round(pct_low,0)`%

+ Percentage Consumed~high~= `r round(pct_high, 0)`%

```{r echo = FALSE, message = FALSE, warning = FALSE}
#MEC of one gallon of gas given multiple SCC values
# MEC = ($/ metric ton) * (metric tons/lb) * (lb/gal)

# $75 SCC
MEC_75 <- 75*(1/2204.62)*(19.6)

# $100
MEC_100 <- 100*(1/2204.62)*(19.6)

# $125
MEC_125 <- 125*(1/2204.62)*(19.6)

#150
MEC_150 <- 150*(1/2204.62)*(19.6)

```

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Find the socially optimum equilibrium price and and quantity

# $75 SCC
MSC_75_Q_eq <- (MEC_75-int_agg)/(m_agg-m_MC)
MSC_75_P_eq <- (m_MC*MSC_75_Q_eq+ MEC_75)

# $100
MSC_100_Q_eq <- (MEC_100-int_agg)/(m_agg-m_MC)
MSC_100_P_eq <- (m_MC*MSC_100_Q_eq+ MEC_100)

# $125
MSC_125_Q_eq <- (MEC_125-int_agg)/(m_agg-m_MC)
MSC_125_P_eq <- (m_MC*MSC_125_Q_eq+ MEC_125)

# $150
MSC_150_Q_eq <- (MEC_150-int_agg)/(m_agg-m_MC)
MSC_150_P_eq <- (m_MC*MSC_150_Q_eq+ MEC_150)

```

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Find the tax revenue generated from each SCC tax level 
# Tax revenue =(gallons of gas)*(MEC)
#Determine proportion returned to two income groups

# $51
#tax rev
tax_rev_51= TED
#tax return
tr_51_low=tax_rev_51*prop_low
tr_51_high=tax_rev_51*prop_high

# $75
#tax revenue
tax_rev_75=MSC_75_Q_eq*MEC_75
#tax return
tr_75_low=tax_rev_75*prop_low
tr_75_high=tax_rev_75*prop_high

# $100
# tax revenue
tax_rev_100=MSC_100_Q_eq*MEC_100
# tax return
tr_100_low=tax_rev_100*prop_low
tr_100_high=tax_rev_100*prop_high

# $125
#tax revenue
tax_rev_125=MSC_125_Q_eq*MEC_125
#tax return
tr_125_low=tax_rev_125*prop_low
tr_125_high=tax_rev_125*prop_high

# $150
#tax revenue
tax_rev_150=MSC_150_Q_eq*MEC_150
#tax return
tr_150_low=tax_rev_150*prop_low
tr_150_high=tax_rev_150*prop_high
```


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Determine quantity purchased at new socially optimal equilibrium 
# $75
MSC_75_Q_eq_high= (MSC_75_P_eq-int_high)/(m_high)

# $100
MSC_100_Q_eq_high= (MSC_100_P_eq-int_high)/(m_high)

# $125
MSC_125_Q_eq_high= (MSC_125_P_eq-int_high)/(m_high)

# $150
MSC_150_Q_eq_high= (MSC_150_P_eq-int_high)/(m_high)

```

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Determine total welfare by finding the total consumer surplus to the group, subtracting the cost of gas, and adding the amount of tax returned to the group

#Consumer surplus: Integrate under the demand curve from 0 to the new quantity purchased at the socially optimum price found above
#Remove the cost to purchase the gas and the add the proportional amount of tax returned

# $51
CS_high_MSC_tax <- integrate(D_high, lower = 0, upper = MSC_Q_eq_high)
wf_high_MSC <- CS_high_MSC$value - MSC_P_eq*MSC_Q_eq_high + tr_51_high

# $75
CS_high_MSC_75 <- integrate(D_high, lower = 0, upper = MSC_75_Q_eq_high)
wf_high_MSC_75<- CS_high_MSC_75$value - (MSC_75_P_eq*MSC_75_Q_eq_high) +tr_75_high

# $100
CS_high_MSC_100 <- integrate(D_high, lower = 0, upper = MSC_100_Q_eq_high)
wf_high_MSC_100<- CS_high_MSC_100$value - (MSC_100_P_eq*MSC_100_Q_eq_high)+ tr_100_high

# $125
CS_high_MSC_125 <- integrate(D_high, lower = 0, upper = MSC_125_Q_eq_high)
wf_high_MSC_125<- CS_high_MSC_125$value - (MSC_125_P_eq*MSC_125_Q_eq_high) + tr_125_high

# $150
CS_high_MSC_150 <- integrate(D_high, lower = 0, upper = MSC_150_Q_eq_high)
wf_high_MSC_150 <- CS_high_MSC_150$value - (MSC_150_P_eq*MSC_150_Q_eq_high) + tr_150_high
```


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Determine quantity purchased at new socially optimal equilibrium

# $75
MSC_75_Q_eq_low= (MSC_75_P_eq-int_low)/(m_low)

# $100
MSC_100_Q_eq_low= (MSC_100_P_eq-int_low)/(m_low)

# $125
MSC_125_Q_eq_low= (MSC_125_P_eq-int_low)/(m_low)

# $150
MSC_150_Q_eq_low= (MSC_150_P_eq-int_low)/(m_low)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Determine total welfare by finding the total consumer surplus to the group, subtracting the cost of gas and environmental damages, and adding the amount of tax returned to the group

#Consumer surplus: Integrate under the demand curve from 0 to the new quantity purchased at the socially optimum price found above

#Remove the cost to purchase the gas, the cost of the environmental damage, and the add the proportional amount of tax returned to the income group

# $51
CS_low_MSC_tax <- integrate(D_low, lower = 0, upper = MSC_Q_eq_low)
wf_low_MSC <- CS_low_MSC$value - (MSC_P_eq*MSC_Q_eq_low) + tr_51_low -tax_rev_51

# $75
CS_low_MSC_75 <- integrate(D_low, lower = 0, upper = MSC_75_Q_eq_low)
wf_low_MSC_75<- CS_low_MSC_75$value - (MSC_75_P_eq*MSC_75_Q_eq_low) +tr_75_low -tax_rev_75

# $100
CS_low_MSC_100 <- integrate(D_low, lower = 0, upper = MSC_100_Q_eq_low)
wf_low_MSC_100<- CS_low_MSC_100$value - (MSC_100_P_eq*MSC_100_Q_eq_low)+ tr_100_low -tax_rev_100

# $125
CS_low_MSC_125 <- integrate(D_low, lower = 0, upper = MSC_125_Q_eq_low)
wf_low_MSC_125<- CS_low_MSC_125$value - (MSC_125_P_eq*MSC_125_Q_eq_low) + tr_125_low -tax_rev_125

# $150
CS_low_MSC_150 <- integrate(D_low, lower = 0, upper = MSC_150_Q_eq_low)
wf_low_MSC_150 <- CS_low_MSC_150$value - (MSC_150_P_eq*MSC_150_Q_eq_low) + tr_150_low -tax_rev_150

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
#Amount of gas sold 

# $51
income_51 <- MSC_P_eq*MSC_Q_eq - tax_rev_51
cost_51 <- integrate (MC, lower=0, upper = MSC_Q_eq)
revenue_51 <- income_51 - cost_51$value

# $75
income_75 <- MSC_75_P_eq*MSC_75_Q_eq - tax_rev_75
cost_75 <- integrate (MC, lower=0, upper = MSC_75_Q_eq)
revenue_75 <- income_75 - cost_75$value

# $100
income_100 <- MSC_100_P_eq*MSC_100_Q_eq - tax_rev_100
cost_100 <- integrate (MC, lower=0, upper = MSC_100_Q_eq)
revenue_100 <- income_100 - cost_100$value

# $125
income_125 <- MSC_125_P_eq*MSC_125_Q_eq - tax_rev_125
cost_125 <- integrate (MC, lower=0, upper = MSC_125_Q_eq)
revenue_125 <- income_125 - cost_125$value

# $150
income_150 <- MSC_150_P_eq*MSC_150_Q_eq - tax_rev_150
cost_150 <- integrate (MC, lower=0, upper = MSC_150_Q_eq)
revenue_150 <- income_150 - cost_150$value

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, include = FALSE}

# Graph for reference
ggplot(data.frame(Q = c(0, 800000)), aes(x = Q)) +
  stat_function(fun = D_agg, aes(color = "Aggregate Demand")) +
  stat_function(fun = D_low, aes(color = "Low Demand")) +
  stat_function(fun = D_high, aes(color = "High Demand")) +
  stat_function(fun = MC, aes(color = "Marginal Cost")) +
  geom_segment(aes(x = 0, y = P_eq, xend = Q_eq, yend = P_eq, color = "Equilibrium Pre-Tax"), linetype = 2)+
  geom_segment(aes(x = Q_eq, y = 0, xend = Q_eq, yend = P_eq, color = "Equilibrium Pre-Tax"), linetype = 2)+
  stat_function(fun = MSC, aes(color = "MSC - 51")) +
  geom_segment(aes(x = 0, y = MSC_P_eq, xend = MSC_Q_eq, yend = MSC_P_eq, color = "MSC - 51"), linetype = 2)+
  geom_segment(aes(x = MSC_Q_eq, y = 0, xend = MSC_Q_eq, yend = MSC_P_eq, color = "MSC - 51"), linetype = 2)+
  geom_segment(aes(x = 0, y = MEC, xend = MSC_Q_eq, yend = MEC, color = "MSC - 51"), linetype = 2)+
# SCC = 75
  geom_segment(aes(x = 0, y = MSC_75_P_eq, xend = MSC_75_Q_eq, yend = MSC_75_P_eq, color = "MSC - 75"), linetype = 2)+
  geom_segment(aes(x = MSC_75_Q_eq, y = 0, xend = MSC_75_Q_eq, yend = MSC_75_P_eq, color = "MSC - 75"), linetype = 2)+
  geom_segment(aes(x = 0, y = MEC_75, xend = MSC_75_Q_eq, yend = MEC_75, color = "MSC - 75"), linetype = 2)+
# SCC = 100
  geom_segment(aes(x = 0, y = MSC_100_P_eq, xend = MSC_100_Q_eq, yend = MSC_100_P_eq, color = "MSC - 100"), linetype = 2)+
  geom_segment(aes(x = MSC_100_Q_eq, y = 0, xend = MSC_100_Q_eq, yend = MSC_100_P_eq, color = "MSC - 100"), linetype = 2)+
  geom_segment(aes(x = 0, y = MEC_100, xend = MSC_100_Q_eq, yend = MEC_100, color = "MSC - 100"), linetype = 2)+
# SCC = 125
  geom_segment(aes(x = 0, y = MSC_125_P_eq, xend = MSC_125_Q_eq, yend = MSC_125_P_eq, color = "MSC - 125"), linetype = 2)+
  geom_segment(aes(x = MSC_125_Q_eq, y = 0, xend = MSC_125_Q_eq, yend = MSC_125_P_eq, color = "MSC - 125"), linetype = 2)+
  geom_segment(aes(x = 0, y = MEC_125, xend = MSC_125_Q_eq, yend = MEC_125, color = "MSC - 125"), linetype = 2)+
# SCC = 150
  geom_segment(aes(x = 0, y = MSC_150_P_eq, xend = MSC_150_Q_eq, yend = MSC_150_P_eq, color = "MSC - 150"), linetype = 2)+
  geom_segment(aes(x = MSC_150_Q_eq, y = 0, xend = MSC_150_Q_eq, yend = MSC_150_P_eq, color = "MSC - 150"), linetype = 2)+
  geom_segment(aes(x = 0, y = MEC_150, xend = MSC_150_Q_eq, yend = MEC_150, color = "MSC - 150"), linetype = 2)+
  scale_x_continuous(name = "Q (Gallons)") +
  scale_y_continuous(name = "P ($/gal)",limits = c(0, NA)) +
  labs(color = "Curve") +
  theme_minimal()

```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Let's make a table!

# Column of SCC:
SCC <- c(51, 75, 100, 125, 150)

# Column of MEC at SCC:
MEC_col <- c(MEC, MEC_75, MEC_100, MEC_125, MEC_150)

# Column of Peq at SCC:
Peq_col <- c(MSC_P_eq, MSC_75_P_eq, MSC_100_P_eq, MSC_125_P_eq, MSC_150_P_eq)

# Column of Qeq at SCC:
Qeq_col <- c(MSC_Q_eq, MSC_75_Q_eq, MSC_100_Q_eq, MSC_125_Q_eq, MSC_150_Q_eq)

# Column of Q high
Qeq_high_col <- c(MSC_Q_eq_high, MSC_75_Q_eq_high, MSC_100_Q_eq_high, MSC_125_Q_eq_high, MSC_150_Q_eq_high)

# Column of Q low
Qeq_low_col <- c(MSC_Q_eq_low, MSC_75_Q_eq_low, MSC_100_Q_eq_low, MSC_125_Q_eq_low, MSC_150_Q_eq_low)

# Column of env burden
TED_col <- c(tax_rev_51, tax_rev_75, tax_rev_100, tax_rev_125, tax_rev_150)

# Column of welfare before tax distribution high
wf_pre_tax_high_col <- c(CS_high_MSC$value - MSC_P_eq*MSC_Q_eq_high, 
                     CS_high_MSC_75$value - (MSC_75_P_eq*MSC_75_Q_eq_high),
                     CS_high_MSC_100$value - (MSC_100_P_eq*MSC_100_Q_eq_high),
                     CS_high_MSC_125$value - (MSC_125_P_eq*MSC_125_Q_eq_high),
                     CS_high_MSC_150$value - (MSC_150_P_eq*MSC_150_Q_eq_high))

# Column of welfare before tax distribution and env burden low
wf_pre_tax_low_col <- c(CS_low_MSC$value - MSC_P_eq*MSC_Q_eq_low, 
                     CS_low_MSC_75$value - (MSC_75_P_eq*MSC_75_Q_eq_low),
                     CS_low_MSC_100$value - (MSC_100_P_eq*MSC_100_Q_eq_low),
                     CS_low_MSC_125$value - (MSC_125_P_eq*MSC_125_Q_eq_low),
                     CS_low_MSC_150$value - (MSC_150_P_eq*MSC_150_Q_eq_low))

# Column of welfare before tax distribution but including env burden low
wf_pre_tax_low_burden_col <- c(CS_low_MSC$value - MSC_P_eq*MSC_Q_eq_low - tax_rev_51, 
                     CS_low_MSC_75$value - (MSC_75_P_eq*MSC_75_Q_eq_low) - tax_rev_75,
                     CS_low_MSC_100$value - (MSC_100_P_eq*MSC_100_Q_eq_low) - tax_rev_100,
                     CS_low_MSC_125$value - (MSC_125_P_eq*MSC_125_Q_eq_low) - tax_rev_125,
                     CS_low_MSC_150$value - (MSC_150_P_eq*MSC_150_Q_eq_low)- tax_rev_150)

# column of tax revenue redistributed to high
tr_high_col <- c(tr_51_high, tr_75_high, tr_100_high, tr_125_high, tr_150_high)

# column of tax revenue redistributed to low
tr_low_col <- c(tr_51_low, tr_75_low, tr_100_low, tr_125_low, tr_150_low)

# Column of welfare after tax distribution high
welfare_high_col <- c(wf_high_MSC, wf_high_MSC_75, wf_high_MSC_100, wf_high_MSC_125, wf_high_MSC_150)

# Column of welfare after tax distribution low (includes env burden)
welfare_low_col <- c(wf_low_MSC, wf_low_MSC_75, wf_low_MSC_100, wf_low_MSC_125, wf_low_MSC_150)

# Column of gas producer welfare
gas_rev_col <- c(revenue_51, revenue_75, revenue_100, revenue_125, revenue_150)

# Table
part_5_table <- data.frame(SCC = SCC, MEC = MEC_col, Peq = Peq_col, Qeq = Qeq_col, Qeq_high = Qeq_high_col, Qeq_low = Qeq_low_col, TED = TED_col,
                           wf_pre_tax_high = wf_pre_tax_high_col,
                           wf_pre_tax_low = wf_pre_tax_low_col,
                           wf_pre_tax_low_burden = wf_pre_tax_low_burden_col,
                           tr_high = tr_high_col, tr_low = tr_low_col,
                           welfare_high = welfare_high_col, 
                           welfare_low = welfare_low_col,
                           gas_rev = gas_rev_col)
```


First, the range of SCC values were converted to marginal external cost (MEC) of one gallon of gas using the same formula provided in Part 1 above. The MEC for each SCC value equals the SCC-based gasoline tax:

+ $51/ton CO~2~ results in an MEC/SCC-based gasoline tax of \$`r round(MEC,2)`/gal
+ $75/ton CO~2~ results in an MEC/SCC-based gasoline tax of \$`r round(MEC_75,2)`/gal
+ $100/ton CO~2~ results in an MEC/SCC-based gasoline tax of \$`r round(MEC_100,2)`/gal
+ $125/ton CO~2~ results in an MEC/SCC-based gasoline tax of \$`r round(MEC_125,2)`/gal
+ $150/ton CO~2~ results in an MEC/SCC-based gasoline tax of \$`r round(MEC_150,2)`/gal

Second, the new marginal social cost curve for each level of tax was determined to be the Marginal Cost curve determined in Part 2 plus the SCC-based gasoline tax. For example, MSC~[SCC = \$100]~ = `r round(m_MC,6)`Q + `r round(MEC_100,2)`

Third, the socially optimal equilibrium quantity of gas sold was calculated by finding the intersection of the aggregate demand curve from Part 2 and the MSC.

Fourth, the total tax revenue generated by the tax was calculated by multiplying the socially optimal equilibrium quantity of gas sold that was determined above by the MEC. Then, the amount of tax revenue redistributed to each income group was calculated by multiplying the total tax revenue generated by the original "Percentage Consumed" of each income group determined above.

Fifth, the socially optimal price of gasoline was calculated by solving the total aggregate demand curve for P~eq~ given the Q~eq~ determined above.

Sixth, the quantity of gas purchased by each income group was determined by solving each group's demand curve (determined in Part 2) given the P~eq~ determined above.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Create stylized table for report of just Ps and Qs

part_5_table %>% 
  select(SCC, Peq, Qeq, Qeq_high, Qeq_low) %>% 
  kable(digits = 2, format.args = list(big.mark = ","), 
    caption = "Table 1",
    col.names = c("SCC (USD/ton CO~2~)", "Price of gas/gal", "Total Q of gas sold (gal)", "Q of gas sold (gal) to high demanders", "Q of gas sold (gal) to low demanders")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = TRUE)
```


Seventh, welfare of each consumer group was determined by integrating each group's demand curve from 0 to the quantity of gas consumed by the group given the MSC/SCC-based gasoline tax. Then the cost of purchasing the gas (P~eq~*Q~eq~) was subtracted from the integral value. For the low income group, total environmental damage  was also subtracted from the integral. Lastly, tax revenue redistributed to each group was added. Results are shown in the table below.

Eighth, welfare to gas producers was calculated by multiplying the price of gas by the quantity sold and subtracting the tax revenue and the cost of production. Cost of production was determined as the integral of the marginal cost curve from 0 to the equilibrium quantity sold.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Create stylized table for report

part_5_final <- part_5_table %>% 
  select(SCC, tr_high, welfare_high, tr_low, TED, welfare_low, gas_rev)

part_5_final %>% 
  kable(digits = 0, format.args = list(big.mark = ","), 
    caption = "Table 2 (all values in USD unless otherwise noted)",
    col.names = c("SCC (USD/ton CO~2~)", "Tax revenue to high demanders", "Welfare to high demanders", "Tax revenue to low demanders", "Total Environmental Damage", "Welfare to low demanders","Welfare to gas producers")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = TRUE)
```

Notably, increased values of the SCC increases the value of total environmental damage. Because the low income group bears the entirety of the cost of the environmental damage, their welfare continuously decreases with increasing SCC. Even though the group receives `r round(pct_low,0)`% of the tax revenue generated as well as surplus from gas purchased, this is not enough to negate the cost of the environmental damage. Ultimately, at an SCC of \$125 and \$150, the welfare to the low income group is negative, indicating that the group experiences costs, rather than welfare, with this tax scheme. High income group welfare increases slightly with the increasing SCC while welfare to gas producers decreases slightly.


# Part 6

*Suppose a new electric car technology is invented but only the “high” income group has access to these expensive new cars. This lowers the gasoline demand curve for the “high” income group by half (vertically).*
 
Begin by deriving a new demand curve for the high income group. 

Solve for a new y-intercept for the demand curve of high demanders by reducing the previous y-intercept (P = `r round(int_high,2)`) by half. The slope for the new demand curve will remain the same (m = `r round(m_high,6)`) because elasticity of high demand has not changed, only the number of high demanders for gasoline has shifted.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Determine the new demand curve for the high income group
#the y-intercept is lowered by half

int_high_e <- (int_high/2)
D_high_e<- function(Q){m_high*Q + int_high_e}
```
 
The resulting new high income demand curve is D~high-e~ = `r round(m_high,6)`*Q + `r round(int_high_e,2)` 

Use the new high income demand curve to generate a new aggregate demand curve by horizontally summing individual demand curves of low income group (unchanged) and new high income group demand. To sum horizontally, solve each equation for Q, add both equations together, and re-solve for P. The new high income demand curve is lower than the low income demand curve. This means that the aggregate demand equals the demand for the low income consumer group at prices where only the low demand consumer group has demand (i.e. the "kinked" portion of the new aggregate demand curve). The resulting aggregate demand curve when electric technology is introduced is therefore:

```{r echo = FALSE, message = FALSE, warning = FALSE}
#generate a new aggregate demand curve by summing horizontally

# Solve for intercepts for aggregate demand
int_agg_e<- ((int_low/m_low)+(int_high_e/m_high))/(1/(m_low)+1/(m_high))

# Create functions for all three demand curves
D_agg_e<- function(Q){m_agg*Q + int_agg_e}
```


D~agg-e~ , P < `r round(int_high_e,1)` = `r round(m_agg,6)`*Q + `r round(int_agg_e,2)`

D~agg-e~ , P > `r round(int_high_e,1)` = `r round(m_low,6)`Q + `r round(int_low,1)`

```{r echo = FALSE, message = FALSE, warning = FALSE, include = FALSE}
# Graph for reference
ggplot(data.frame(Q = c(0, 800000)), aes(x = Q)) +
  stat_function(fun = D_agg_e, aes(color = "Agg.Demand-Electric")) +
  stat_function(fun = D_low, aes(color = "Low Demand")) +
  stat_function(fun = D_high_e, aes(color = "High Demand-Electric")) +
  stat_function(fun = MC, aes(color = "Supply")) +
   stat_function(fun = D_high, aes(color = "High Demand")) +
   stat_function(fun = D_agg, aes(color = "Agg. Demand")) +
  scale_x_continuous(name = "Q (Gallons)") +
  scale_y_continuous(name = "P ($/gal)",limits = c(0, NA)) +
  labs(color = "Curve") +
  theme_minimal()
```


#### (a) What is total gas consumption? 

The new equilibrium quantity of gas is found by setting the aggregate demand curve with electric technology equal to the supply curve (unchanged from earlier parts) and solving for Q.

`r round(m_agg,6)`* Q + `r round(int_agg_e,2)` =  `r round(m_MC,6)`*Q

```{r echo = FALSE, message = FALSE, warning = FALSE}
#solve for the new equilibrium gas quantity sold
# D: P=m1*Q+a
# S: P=m2*Q+b
# m1*Q+a=m2*Q+b
# (m1-m2)*Q=(b-a)
# Q=(b-a)/(m1-m2)

Q_eq_e <- (0-int_agg_e)/(m_agg-m_MC)
```

Q~eq,e~ = `r round(Q_eq_e, 0)` gallons

```{r echo = FALSE, message = FALSE, warning = FALSE}
# Find the new equilibrium price 

# Evaluate the demand curve at the new equilibrium quantity 

P_eq_e <- (m_MC)*(Q_eq_e)
```


#### (b) What is the total environmental externality?

Calculate the total environmental damage assuming the social cost of carbon is $51/ton of CO~2~ emitted. 

Total environmental damage is calculated by multiplying the marginal external cost (MEC = \$`r round(MEC, 2)`/gal) calculated in section 2 by the new equilibrium quantity of gas sold (Q = `r round(Q_eq_e, 0)` gal)

```{r echo = FALSE, message = FALSE, warning = FALSE}
# TED = MEC*Q

TED_e <- MEC*Q_eq_e
```

Total Environmental Damage = \$`r round(TED_e,2)`

#### (c) What value of the gasoline tax makes the total environmental damage the same as the electric car technology?

The total environmental damage is found by multiplying the marginal external cost by the quantity of gas at equilibrium. 

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Find that would result in the same TED
# TED=Q*t
# t= P-3
# Solve for Q of the total marginal cost curve: P=m_MC*Q+t
# Q=(P-t)/m_MC= (P-(P-3))m_MC = 3/m_MC
# TED= (3/m_MC)(P-3)
# Solve for P: P= TED*(m_MC/3)+3
P_e_tax <- ((TED_e*m_MC)/3)+3

#calculate tax: P= 3+t
tax_e<- P_e_tax-3

#Check work 
Q_tax_e <- (P_e_tax-tax_e)/(m_MC)
TED_t_e <- Q_tax_e*tax_e
```

The tax to generate the same total environmental damage as adopting electric vehicle technology is \$`r round(tax_e,2)`. 

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Solve for P at old D aggregate given new Q:
D_agg_q_eq_e <- D_agg(Q_eq_e)

S_q_eq_e <- MC(Q_eq_e)

new_tax <- D_agg_q_eq_e - S_q_eq_e
new_tax

```

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Graph for reference
ggplot(data.frame(Q = c(0, 800000)), aes(x = Q)) +
  stat_function(fun = D_agg_e, aes(color = "Agg Demand Electric")) +
  #stat_function(fun = D_low, aes(color = "Low Demand")) +
  #stat_function(fun = D_high_e, aes(color = "High Demand-Electric")) +
  stat_function(fun = MC, aes(color = "Supply")) +
  #stat_function(fun = D_high, aes(color = "High Demand")) +
  stat_function(fun = D_agg, aes(color = "Agg Demand Original")) +
  
  geom_segment(aes(x = 0, y = S_q_eq_e, xend = Q_eq_e, yend = S_q_eq_e), linetype = 2)+
  geom_segment(aes(x = Q_eq_e, y = 0, xend = Q_eq_e, yend = S_q_eq_e), linetype = 2)+
  
  geom_segment(aes(x = 0, y = D_agg_q_eq_e, xend = Q_eq_e, yend = D_agg_q_eq_e), linetype = 2)+
  geom_segment(aes(x = Q_eq_e, y = S_q_eq_e, xend = Q_eq_e, yend = D_agg_q_eq_e, color = "Optimal tax"), linetype = 1, size = 1.5)+
  
  geom_rect(aes(xmin=0, xmax=Q_eq_e, ymin=0, ymax=MEC, fill="Total Environmental Damage"), alpha=0.3) +
  
  scale_x_continuous(name = "Q (Gallons)") +
  scale_y_continuous(name = "P ($/gal)",limits = c(0, NA)) +
  labs(color = "Curve", fill = "Shading") +
  theme_minimal()

```










